Sampling Bayesian probabilities given only sampled priors

### 2024-11-20 main.tex
Fix the inversion issue, add code

If we next define the Jacobian matrix
\begin{equation}
  \jac \equiv \left.\frac{\partial\hatc}{\partial\vecn}\right|_{\vecq_0, \vecn_0}
\label{eq:jacobian}
\end{equation}
and assume that a first-order Taylor expansion about the nominal values is valid across the domain of interest, then we obtain
\begin{equation}
  \left\langle \Delta\chi^2 \right\rangle \approx \frac{1}{N_\alpha} \sum_\alpha
  \left[ \vecn_\alpha - \hat{\vecn}(u_\alpha)\right]^T \jac^T \covm_c^{-1} \jac
  \left[ \vecn_\alpha - \hat{\vecn}(u_\alpha)\right].
  \label{eq:linearized}
\end{equation}
We ask what is the optimal \emph{linear} model $\hat{\vecn} = \matU \cdot \vecu$ to use.
It is useful to define the matrix \matD\ and its ``square root'' $\DD$ by
\begin{equation}
   \matD \equiv \jac^T \covm_c^{-1} \jac =  \DD^T \DD.
 \end{equation}
Since \matD\ is positive-definite and symmetric, $\DD$ will exist and be real-valued.
If \vecn\ has $N$ elements, and we want to compress the representation
to a vector \vecu\ with $M$ elements, we will use the $M\times N$ projection
matrix $\proj_M$ whose elements are
\begin{equation}
  \left(\proj_M\right)_{ij} \equiv
\begin{cases}
                                            1,  &  i=j\le M \\
                                            0,  & \text{otherwise.}
\end{cases}
\end{equation}
such that $\proj_M \proj_M^T=\ident_M,$ the $M\times M$ identity
matrix. We will assume that \matU\ satisfies
\begin{equation}
  \DD \matU = \matV_U \proj_M^T
\label{eq:vupm}
\end{equation}
for some unitary matrix $\matV_U$.  We can do this because, for any $\matU$, we have the
singular-value decomposition $\DD \matU = \matV_U P_M^T \Lambda_U
W_U^T,$ where $\Lambda_U$ is diagonal, $W_U$ is unitary, and both are
$M\times M$.  We can always absorb these last two factors into the \vecu\ values
to conserve the values of $\hat{\vecn} = \matU \vecu$ while making
\eqq{eq:vupm} hold.

@@

  \item The \matU\ matrix containing the $M$ basis vectors for
    reconstructing \vecn\ is
    \begin{equation}
      \matU = \DD^{-1} \matV_A \proj_M^T.
    \end{equation}
    This requires that $\DD$ be invertible, but if we recall that
    $\matV_A$ are the eigenvectors of $\DD \covm_n \DD,$ we can see
    that we can get a well-defined result as long as the rank of
    \matD\ is $\ge M$.  ???

  \item The compressed $\vecu_\alpha$ for a chosen sample of \vecn\ is
    \begin{equation}
      \vecu_\alpha = \proj_M \matV_A^T \DD \vecn_\alpha,
    \end{equation}
    and the covariance matrix of the compressed samples is
    \begin{equation}
      \covm_u = \left\langle \vecu \vecu^T \right\rangle = \proj_M
      \matV_A^T \DD \covm_n \DD^T \proj_M^T = \text{Diag}(\lambda^A_1,
      \lambda^A_2, \ldots \lambda^A_M).
    \end{equation}
The components of the compressed \vecu\ coefficients are therefore
uncorrelated.  Component $u_i$ has a variance
$\lambda^A_i,$ which is also equal to the reduction in $\left\langle
  \Delta \chi^2 \right\rangle$ that we will accrue if we include basis
vector $\mathbf{U}_i$ into our linear reconstruction of \vecn.
\end{itemize}

>>>>

If we next define the Jacobian matrix
\begin{equation}
  \jac \equiv \left.\frac{\partial\hatc}{\partial\vecn}\right|_{\vecq_0, \vecn_0}
\label{eq:jacobian}
\end{equation}
and assume that a first-order Taylor expansion about the nominal values is valid across the domain of interest, then we obtain
\begin{equation}
  \left\langle \Delta\chi^2 \right\rangle \approx \frac{1}{N_\alpha} \sum_\alpha
  \left[ \vecn_\alpha - \hat{\vecn}(u_\alpha)\right]^T \jac^T \covm_c^{-1} \jac
  \left[ \vecn_\alpha - \hat{\vecn}(u_\alpha)\right].
  \label{eq:linearized}
\end{equation}
We ask what is the optimal \emph{linear} model $\hat{\vecn} = \matU \cdot \vecu$ to use.
It is useful to define the matrix \matD, its eigensystem
$(\matV_D,\Lambda_D),$  \ and its ``square root'' $\DD$ by
\begin{equation}
   \matD \equiv \jac^T \covm_c^{-1} \jac =  \matV_D \Lambda_D
   \matV_D^T = \left(\matV_D \Lambda_D^{1/2}\right)
   \left(\Lambda_D^{1/2}\matV_D^T \right) = \DD^T \DD.
 \end{equation}
Since \matD\ is non-negative and symmetric, $\DD$ will exist and be
real-valued with non-negative singular values.
If \vecn\ has $N$ elements, and we want to compress the representation
to a vector \vecu\ with $M$ elements, we will use the $M\times N$ projection
matrix $\proj_M$ whose elements are
\begin{equation}
  \left(\proj_M\right)_{ij} \equiv
\begin{cases}
                                            1,  &  i=j\le M \\
                                            0,  & \text{otherwise.}
\end{cases}
\end{equation}
such that $\proj_M \proj_M^T=\ident_M,$ the $M\times M$ identity
matrix. We will assume that \matU\ satisfies
\begin{equation}
  \DD \matU = \matV_U \proj_M^T
\label{eq:vupm}
\end{equation}
for some unitary matrix $\matV_U$.  We can do this because, for any $\matU$, we have the
singular-value decomposition $\DD \matU = \matV_U P_M^T \Lambda_U
W_U^T,$ where $\Lambda_U$ is diagonal, $W_U$ is unitary, and both are
$M\times M$.  We can always absorb these last two factors into the \vecu\ values
to conserve the values of $\hat{\vecn} = \matU \vecu$ while making
\eqq{eq:vupm} hold.

@@

  \item The \matU\ matrix containing the $M$ basis vectors for
    reconstructing \vecn\ is
    \begin{equation}
      \matU = \DD^{-1} \matV_A \proj_M^T = \matV_D \Lambda_D^{-1/2}  \matV_A \proj_M^T.
    \end{equation}
    In inverting the (non-negative, diagonal) matrix $\Lambda_D^{1/2},$ one should null the
    inverses of elements that are $\lesssim 10^{-3}$ of the largest
    element, to avoid adding large fluctuations to \matU\ that address
    noise or roundoff-error terms that contribute negligibly to $\chi^2.$

  \item The compressed $\vecu_\alpha$ for a chosen sample of \vecn\ is
    \begin{equation}
      \vecu_\alpha = \proj_M \matV_A^T \DD \vecn_\alpha = \proj_M
      \matV_A^T \Lambda_D^{1/2} \matV_D^T \vecn_\alpha,
    \end{equation}
    and the covariance matrix of the compressed samples is
    \begin{equation}
      \covm_u = \left\langle \vecu \vecu^T \right\rangle = \proj_M
      \matV_A^T \DD \covm_n \DD^T \matV_A \proj_M^T = \text{Diag}(\lambda^A_1,
      \lambda^A_2, \ldots \lambda^A_M).
    \end{equation}
The components of the compressed \vecu\ coefficients are therefore
uncorrelated.  Component $u_i$ has a variance
$\lambda^A_i,$ which is also equal to the reduction in $\left\langle
  \Delta \chi^2 \right\rangle$ that we will accrue if we include basis
vector $\mathbf{U}_i$ into our linear reconstruction of \vecn.
\end{itemize}
The compression thus consists of choosing the desired number of rows
of the matrix $ \matV_A^T \Lambda_D^{1/2} \matV_D^T$, and the modes
are the corresponding columns of this matrix's inverse, with each
included mode reducing the mean compression error in mean $\chi^2$ by $\lambda^A_i.$
